---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Code for A new car-following model with incorporation of Markkula's framework of sensorimotor control in sustained motion tasks

This code repository contains functions to use the `Markkula Intelligent Driver Model (MIDM)` developed in the [paper titled above](https://www.sciencedirect.com/science/article/abs/pii/S0191261524000936). There are two code files in the `code` folder:  

* `DoControlModelTimeStep_new.R`: This is essentially the same functionality in R as written by Gustav Markkula in MATLAB. The function `DoControlModelTimeStep_new` estimates the intermittent control in a simulation.  
* `simulate_full_IDM_intermittent_new.R`: This uses the `DoControlModelTimeStep_new` in a car-following model for predicting speed, acceleration, jerk, and spacing of a following vehicle.    

```{r, echo=FALSE}
source(here::here("code/DoControlModelTimeStep_new.R"))
source(here::here("code/simulate_full_IDM_intermittent_new.R"))
```

#### Example

Load the example data:
```{r}
dat <- read.csv("data/demo_trajectory_data.csv")
head(dat)
```

Other than vehicle IDs and lane number, all variables are needed as inputs to the MIDM. Lead vehicle and following vehicle variables use the prefixes `LV` and `ED` respectively. 

MIDM requires the complete trajectory of the lead vehicle and only the first values of the following vehicle variables. But the function `simulate_full_IDM_intermittent_new` asks for a dataframe with both `LV` and `ED` variables to later compare the observed trajectories with the MIDM trajectories.  

In this example, we use the following arguments for the model parameters:  
```{r}
model_arguments <- c(
  Tg = 1.368, A = 3.325, b = 4.189, v_0 = 28, small_delta = 5.816,
  s_0 = 1.825, accumulator_gain = 3022.003, A_plus = 25.055, A_minus = -24.579,
  lambda = 0.744, epsilon_0 = 2.404, tau_s = 0.026, tau_m = 0.029,
  sigma_n = 0.395, sigma_m = 0.291, G_Duration = 0.666, G_StdDev = 0.373,
  H_Time_error_starts_disappearing = 0.367, H_Time_error_disappeared = 1.524,
  H_error_pred_sd_to_include = 1.92
)
```

Now we can use the MIDM to generate trajectories:  
```{r}
results <- simulate_full_IDM_intermittent_new(
  dat,
  model_arguments[1], 
  model_arguments[2], 
  model_arguments[3], 
  model_arguments[4], 
  model_arguments[5], 
  model_arguments[6], 
  1,
  model_arguments[7], 
  model_arguments[8], 
  model_arguments[9], 
  model_arguments[10], 
  model_arguments[11], 
  model_arguments[12],   
  model_arguments[13], 
  model_arguments[14], 
  model_arguments[15], 
  model_arguments[16], 
  model_arguments[17], 
  model_arguments[18],
  model_arguments[19],
  model_arguments[20]
)

head(results)
```

The new variables in the `results` are:  

* `ED_pred_speed_mps`: Following vehicle speed predicted by MIDM  
* `ED_pred_acc_mps2`: Following vehicle acceleration predicted by MIDM  
* `LV_pred_frspacing_m`: Spacing (front bumper to rear bumper) predicted by MIDM  
* `ED_pred_jerk_mps3`: Following vehicle jerk predicted by MIDM  

... and others.  

We can now compare the results with oberved data: 
```{r}
library(ggplot2)

ggplot(data = results, aes(x = Time)) +
  geom_line(aes(y = LV_speed_mps, color = "LV Speed")) +
  geom_line(aes(y = ED_speed_mps, color = "ED Speed")) +
  geom_line(aes(y = ED_pred_speed_mps, color = "MIDM Speed")) +
  labs(y = "Speed (m/s)") +
  scale_color_manual(values = c("darkblue", "grey30", "darkgreen")) +
  theme_minimal() 

ggplot(data = results, aes(x = Time)) +
  geom_line(aes(y = LV_acc_mps2, color = "LV Acc.")) +
  geom_line(aes(y = ED_acc_mps2, color = "ED Acc.")) +
  geom_line(aes(y = ED_pred_acc_mps2, color = "MIDM acc")) +
  labs(y = "Acceleration (m/s^2)") +
  scale_color_manual(values = c("darkblue", "grey30", "darkgreen")) +
  theme_minimal() 

ggplot(data = results |> dplyr::filter(Time > 0.2), aes(x = Time)) +
  geom_line(aes(y = ED_jerk_mps3_lag, color = "ED Jerk")) +
  geom_line(aes(y = ED_pred_jerk_mps3, color = "MIDM Jerk")) +
  labs(y = "Jerk (m/s^3)") +
  scale_color_manual(values = c("darkblue", "darkgreen")) +
  theme_minimal() 
```

### Terms of use

> THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS
OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT.
IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY
CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT,
TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE
SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.

You are free to use and modify the code in this repository in your own work only. Reproduction, display, distribution and transmission of the codes to third parties is forbidden. Please abide by these terms:   

* Use the code for academic purposes only.  
* Cite the [Transportation Research Part-B paper](https://www.sciencedirect.com/science/article/abs/pii/S0191261524000936) when you publish any derivative work based on this code/study.
